---
title: "Midterm Paper"
author: "econ4ever"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 

#all libraries go here
library(ezids)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(data.table)
library(vtable)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

## Introduction

We have data over a 26 month period i.e. from January 2020 to the end of February 2022. The URL for the data is https://dcatlas.dcgis.dc.gov/crimecards/. 

```{r}
#import dataset to environment 
crimedata <- read.csv("data/dc-crimes-search-results.csv")

#removing redundant variables
crimedata1 <- crimedata[, c(2,6,9,10,11,12,23,26,27)]

#changing the blank spaces to NAs
crimedata1[crimedata1 == "" | crimedata1 == " "] <- NA

#converting REPORT_DAT from character to date format 
crimedata1$REPORT_DAT <- as.Date(crimedata1$REPORT_DAT, "%m/%d/%Y")

xkablesummary(crimedata1)

```

Let's get actual counts of the crimes committed in each category.

```{r}

#dividing up the years

crimedata20 <- subset(crimedata1[crimedata1$YEAR == 2020, ])
crimedata21 <- subset(crimedata1[crimedata1$YEAR == 2021, ])
crimedata22 <- subset(crimedata1[crimedata1$YEAR == 2022, ])

table(crimedata20$offensekey)

table(crimedata21$offensekey)

table(crimedata22$offensekey)

```

Is there a correlation between time of day and type of crime committed? 

```{r, results='markup'}

crimedata1 %>% ggplot(aes(y = OFFENSE, fill = SHIFT)) +
  geom_histogram(stat = "count", binwidth = 1) + 
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "light gray"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank()) +
    #  axis.text.x = element_text(angle = 90)) +
  
  labs(y = "Type of Offense", x = "Frequency", title = "Crime Occurrance At Different Times of Day", 
      fill = "Time of Day")

chisq.test(crimedata1$SHIFT, crimedata1$OFFENSE)

```

What type of crime is most likely to occur in each ward? 

```{r, results='markup'}

crimedata1 %>% ggplot(aes(x = as.factor(WARD), fill = OFFENSE)) +
  geom_histogram(stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "light gray"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank()) +
  #scale_x_continuous(breaks = c(seq(1,8, by=1),9999)) +
 labs(x = "Ward", fill = "Type of Offense", y = "Frequency", title = "Types of Crimes Common in Different Wards")


```

Is there a correlation between property crime and violent crime? 

```{r}

propcrimes <- subset(crimedata1[crimedata1$offensegroup == "property",])
vicrimes <- subset(crimedata1[crimedata1$offensegroup == "violent",])

```





```{r}
#subsetting the data according to wards

ward1 <- subset(crimedata1[crimedata1$WARD == 1,])

ward2 <- subset(crimedata1[crimedata1$WARD == 2,])

ward3 <- subset(crimedata1[crimedata1$WARD == 3,])

ward4 <- subset(crimedata1[crimedata1$WARD == 4,])

ward5 <- subset(crimedata1[crimedata1$WARD == 5,])

ward6 <- subset(crimedata1[crimedata1$WARD == 6,])

ward7 <- subset(crimedata1[crimedata1$WARD == 7,])

ward8 <- subset(crimedata1[crimedata1$WARD == 8,])

```

Now, let's make a summary of the data.

```{r, results='markup'}
summary1 <- summary(crimedata1)
summary1
table1 <- sumtable(crimedata1)
```

Let's check out crime incidents in the different wards.

```{r, results='markup'}
#creating summary tables of the offense data by ward

table(ward1$offensekey)

table(ward2$offensekey)

table(ward3$offensekey)

table(ward4$offensekey)

table(ward5$offensekey)

table(ward6$offensekey)

table(ward7$offensekey)

table(ward8$offensekey)

```

Now let's check out crime incidents in the different wards by year.

```{r, results='markup'}

ward120 <- subset(ward1[ward1$YEAR == 2020, ])
ward121 <- subset(ward1[ward1$YEAR == 2021, ])

ward220 <- subset(ward2[ward2$YEAR == 2020, ])
ward221 <- subset(ward2[ward2$YEAR == 2021, ])

ward320 <- subset(ward3[ward3$YEAR == 2020, ])
ward321 <- subset(ward3[ward3$YEAR == 2021, ])

ward420 <- subset(ward4[ward4$YEAR == 2020, ])
ward421 <- subset(ward4[ward4$YEAR == 2021, ])

ward520 <- subset(ward5[ward5$YEAR == 2020, ])
ward521 <- subset(ward5[ward5$YEAR == 2021, ])

ward620 <- subset(ward6[ward6$YEAR == 2020, ])
ward621 <- subset(ward6[ward6$YEAR == 2021, ])

ward720 <- subset(ward7[ward7$YEAR == 2020, ])
ward721 <- subset(ward7[ward7$YEAR == 2021, ])

ward820 <- subset(ward8[ward8$YEAR == 2020, ])
ward821 <- subset(ward8[ward8$YEAR == 2021, ])


table(ward120$offensekey)
table(ward121$offensekey)

table(ward220$offensekey)
table(ward221$offensekey)

table(ward320$offensekey)
table(ward321$offensekey)

table(ward420$offensekey)
table(ward421$offensekey)


table(ward520$offensekey)
table(ward521$offensekey)

table(ward620$offensekey)
table(ward621$offensekey)

table(ward720$offensekey)
table(ward721$offensekey)

table(ward820$offensekey)
table(ward821$offensekey)

#saving the tables as dataframes to access/play with data

table821 <- as.data.frame(table(ward821$offensekey))
crimetable21 <- as.data.frame(pivot_wider(table821, names_from = "Var1", values_from =  "Freq"))
crimetable21$Ward <- 8

table721 <- as.data.frame(table(ward721$offensekey))
table7211 <- as.data.frame(pivot_wider(table721, names_from = "Var1", values_from =  "Freq"))
table7211$Ward <- 7

table621 <- as.data.frame(table(ward621$offensekey))
table6211 <- as.data.frame(pivot_wider(table621, names_from = "Var1", values_from =  "Freq"))
table6211$Ward <- 6
table6211$`property|arson` <- 0

table521 <- as.data.frame(table(ward521$offensekey))
table5211 <- as.data.frame(pivot_wider(table521, names_from = "Var1", values_from =  "Freq"))
table5211$Ward <- 5
table5211$`property|arson` <- 0

table421 <- as.data.frame(table(ward421$offensekey))
table4211 <- as.data.frame(pivot_wider(table421, names_from = "Var1", values_from =  "Freq"))
table4211$Ward <- 4
table4211$`property|arson` <- 0

table321 <- as.data.frame(table(ward321$offensekey))
table3211 <- as.data.frame(pivot_wider(table321, names_from = "Var1", values_from =  "Freq"))
table3211$Ward <- 3
table3211$`property|arson` <- 0

table221 <- as.data.frame(table(ward221$offensekey))
table2211 <- as.data.frame(pivot_wider(table221, names_from = "Var1", values_from =  "Freq"))
table2211$Ward <- 2

table121 <- as.data.frame(table(ward121$offensekey))
table1211 <- as.data.frame(pivot_wider(table121, names_from = "Var1", values_from =  "Freq"))
table1211$Ward <- 1

wardcrime21 <- rbind(crimetable21, table1211, table2211, table3211, table4211, table5211, table6211, table7211)

write_csv(wardcrime21, "wardcrime21.csv")

#repeating the process for the year 2020

table820 <- as.data.frame(table(ward820$offensekey))
table8200 <- as.data.frame(pivot_wider(table820, names_from = "Var1", values_from =  "Freq"))
table8200$Ward <- 8
table8200$`property|arson` <- 0

table720 <- as.data.frame(table(ward720$offensekey))
table7200 <- as.data.frame(pivot_wider(table720, names_from = "Var1", values_from =  "Freq"))
table7200$Ward <- 7

table620 <- as.data.frame(table(ward620$offensekey))
table6200 <- as.data.frame(pivot_wider(table620, names_from = "Var1", values_from =  "Freq"))
table6200$Ward <- 6

table520 <- as.data.frame(table(ward520$offensekey))
table5200 <- as.data.frame(pivot_wider(table520, names_from = "Var1", values_from =  "Freq"))
table5200$Ward <- 5

table420 <- as.data.frame(table(ward420$offensekey))
table4200 <- as.data.frame(pivot_wider(table420, names_from = "Var1", values_from =  "Freq"))
table4200$Ward <- 4
table4200$`property|arson` <- 0

table320 <- as.data.frame(table(ward320$offensekey))
table3200 <- as.data.frame(pivot_wider(table320, names_from = "Var1", values_from =  "Freq"))
table3200$Ward <- 3
table3200$`property|arson` <- 0

table220 <- as.data.frame(table(ward220$offensekey))
table2200 <- as.data.frame(pivot_wider(table220, names_from = "Var1", values_from =  "Freq"))
table2200$Ward <- 2

table120 <- as.data.frame(table(ward120$offensekey))
table1200 <- as.data.frame(pivot_wider(table120, names_from = "Var1", values_from =  "Freq"))
table1200$Ward <- 1
table1200$`property|arson` <- 0

wardcrime20 <- rbind(table1200, table2200, table3200, table4200, table5200, table6200, table7200, table8200)

write_csv(wardcrime20, "wardcrime20.csv")

#by = c("Ward", "property|arson", "property|burglary", "property|motor vehicle theft", #"property|theft f/auto", "property|theft/other", "violent|assault w/dangerous weapon", #"violent|homicide", "violent|robbery", "violent|sex abuse")

```


Let's group all violent incidences together and all property incidences together and see if there is a correlation.  

```{r, results='markup'}
wardcrime21$totalproperty <- wardcrime21$`property|burglary`+wardcrime21$`property|motor vehicle theft`+wardcrime21$`property|theft f/auto`+wardcrime21$`property|theft/other`

wardcrime20$totalproperty <- wardcrime20$`property|burglary`+wardcrime20$`property|motor vehicle theft`+wardcrime20$`property|theft f/auto`+wardcrime20$`property|theft/other`

wardcrime21$totalviolent <- wardcrime21$`violent|assault w/dangerous weapon`+wardcrime21$`violent|homicide`+wardcrime21$`violent|robbery`+wardcrime21$`violent|sex abuse`

wardcrime20$totalviolent <- wardcrime20$`violent|assault w/dangerous weapon`+wardcrime20$`violent|homicide`+wardcrime20$`violent|robbery`+wardcrime20$`violent|sex abuse`

cor.test(wardcrime21$totalproperty,wardcrime21$totalviolent)
cor.test(wardcrime20$totalproperty,wardcrime20$totalviolent)
```



Let's map out the wards.

```{r, results='markup'}

ward1 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "light pink"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
       axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,5000, by=500),9999)) +labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 1")


ward2 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "light blue"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
       axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,5000, by=500),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 2")


ward3 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "light green"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
       axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,2000, by=200),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 3")


ward4 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "gold"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
       axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,2000, by=200),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 4")


ward5 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "turquoise"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
       axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,3000, by=300),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 5")


ward6 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "purple"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
       axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,4000, by=400),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 6")


ward7 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "lime green"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
       axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,4000, by=300),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 7")


ward8 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "sky blue"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
       axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,4000, by=300),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 8")

```

What types of offenses are more common at different times of year?

```{r, results='markup'}

monthdata <- as.data.frame(month(as.POSIXlt(crimedata1$REPORT_DAT, format="%m/%d/%Y")))

offensedata <- as.data.frame(as.factor(crimedata1$OFFENSE))

setnames(monthdata, 'month(as.POSIXlt(crimedata1$REPORT_DAT, format = "%m/%d/%Y"))', 'month')
setnames(offensedata, 'as.factor(crimedata1$OFFENSE)', "offense")

wardcrime20 %>% ggplot(aes(y = `property|theft/other`, x = Ward)) +
  geom_col(fill = "white", color = "blue") + 
  geom_text(label=wardcrime20$`property|theft/other`, vjust = 5) +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "sky blue"),
      #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank()) +
     #  axis.text.x = element_text(angle = 90)) +
     scale_x_continuous(breaks = c(seq(1,8, by=1),9999)) +
       labs(x = "Ward", y = "Frequency of Theft", title = "Frequency of Theft in Different Wards for 2020")

xkablesummary(wardcrime20)

xkablesummary(wardcrime21)

ward4 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "gold"),
        #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,2000, by=200),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 4")


ward5 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "turquoise"),
        #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,3000, by=300),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 5")


ward6 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "purple"),
        #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,4000, by=400),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 6")



ward7 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "lime green"),
        #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,4000, by=300),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 7")



ward8 %>% ggplot(aes(x = as.factor(OFFENSE), fill = SHIFT)) +
  geom_histogram(binwidth = 1, stat = "count") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        panel.grid.major.y = element_line(color = "sky blue"),
        #  panel.grid.minor.y = element_line(color = "black"),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(seq(0,4000, by=300),9999)) +
  labs(x = "Type of Offense", fill = "Time", y = "Frequency", title = "Types of Crimes Common at Different Times of Day in Ward 8")
```

Is the method of the crime dependent on the offense type?

```{r, results='markup'}
chisq.test(crimedata1$METHOD,crimedata1$offensegroup)
```

Is there relationship between variables ward and OFFENSE?
```{r, results='markup'}
two_way = table(crimedata1$WARD, crimedata1$OFFENSE)
two_way
prop.table(two_way)#cell percentages
prop.table(two_way,1)#row percentages
prop.table(two_way,2)#column percentages
chisq.test(two_way)
```

Which kind of crime happen most frequently during the day/evening/midnight?
```{r, results='markup'}
two_way1 = table(crimedata1$SHIFT,crimedata1$OFFENSE)
two_way1
prop.table(two_way1,1)
```

