---
title: 'Summary Paper: Crime in DC'
author: "Group 4: Rebecca Felix, Jiaxi Liu, Sophia Mir"
subtitle: 'DATS 6101 - Spring 2022'
# date: "today"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 

#all libraries go here
library(ezids)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(data.table)
library(vtable)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

<br>

__Introduction__ \

As a research topic criminal activity is ripe with curious minds trying to figure out who is at fault. Particularly in the field of economics, as researchers and policymakers want to know how to curtail crime rates – is it through creating more jobs, by increasing incarceration time or fines, or through developing more community supports? As budding economists, we are curious of the same mechanism and wanted to explore the data regarding criminal activity in Washington D.C. We wanted to see what we could unearth using the tools we have learned thus far in this class. As we began looking through the research conducted on crime rates and possible causes for changes in crime rates, we saw a strong theme emerge between crime rates and unemployment based around the seminal work of economist Gary Becker. He was one of the first people who tried to wrangle why people commit crimes and if we could create policy to de-incentivize it. \

Economic theory dictates that as unemployment rises, the opportunity cost of committing a crime falls leading to more incidents of crime. Burdett et al (2003) use the work of Gary Becker (1968) and others to further probe at this idea and create a model more sensitive to the heterogeneity in the labor force. They consider possible previous sentences, wage rates, and they try to predict the likelihood of turning to crime given factors such as age and labor wage rates. While they admit that geographical conditions may play a role as well, their findings show that for the most part good labor opportunities with reasonable wages do result in a lower crime rate. \

Janko & Popli (2015) similarly build upon the work of Cantor and Land (1985) who created a theoretical framework to explain the relationship between unemployment and crime. The framework lies on two foundational ideas: opportunity and motivation. With the former being that as economic activity falls so do the opportunities for property crimes and the latter that as economic activity falls the incentives for committing crime increase. On testing this framework using an econometric model, Janko & Popli found that after they accounted for types of crimes and regional differences, there seemed to be no significant relationship between economic activity and crime. \

Frederick & Jozefowicz (2018) created a different econometric model to test Cantor and Land’s theoretical framework in rural and urban Pennsylvania. While their findings were similar to Janko & Popli’s (2015) regarding the rural sectors of Pennsylvania, they found a positive relationship between unemployment and crime in the urban sectors. Edmark (2005) ran a fixed-effects study using panel data from Swedish countries from the years 1988-1999, a time period in which unemployment rose from 1.6% to 10.4%. Unlike the inconsistencies that Janko & Popli (2015) and Frederick & Jozefowicz (2018) found, Edmark’s model showed a positive and significant relationship between property crimes and unemployment rates in the Swedish counties studied.  \

Lee (2016) examined the mixed results of previous studies and the inconsistency between theory and empirical findings and he decided to add more elements to the model – returns on crime, apprehension rates and unemployment benefits. He found that between those three variables the apprehension rate was the biggest predictor in the direction and significance of the relationship between unemployment and crime. \

These are the theories and research that our exploration of the data is based on. For this initial paper we are exploring the first half of the crime-unemployment dichotomy and we will add the unemployment data at the model-building stage.

__Data Source and Limitations__ \

To examine the crime rates in Washington DC (DC) we used incident data gathered by the Metropolitan Police Department (MPD) which we accessed from their website. The data had information on over 50,000 recorded offenses which were reported to, and investigated by, the MPD. The dataset had incidents ranging from January 1st, 2020, through February 28th, 2022. There was a total of 29 variables made up of qualitative data in the dataset but only around five variables had complete and unique data. The dataset classified all crimes into two larger groups of property crimes and violent crimes and further into nine subcategories which we will talk about. \

While it did have truly interesting data, there were a couple of limitations with using this dataset. The first one was that it is not truly representative of crime in the District of Columbia as this crime database system relies on self-report which can lead to underrepresentation of certain crimes and overrepresentation of others. Our other limitation was that because this initial dataset consisted of primarily nominal data, it was difficult to run meaningful statistical tests on it in our exploratory data analysis. As you will see below this led us to largely running chi-squared tests to test for relationships among the different variables given in the data and those we isolated from the data. \

While we did aggregate the incidences by time and by geographical location, running normality tests on it would have given us meaningless results. Running non-parametric tests on the data was also impossible as once aggregated by area, time, types of crimes, there were no categories which had an equal number of observations (n) which could be compared meaningfully. We considered taking random samples of equal n from the different geographical areas (wards) to run a nonparametric comparison of means (Wilcoxon test), but it would not have told us anything of significance that we could not learn through the data summaries. \

As we have already established that we did not believe this dataset to include the entirety of criminal activity occurring in the DC area, we do not assume that our data is normally distributed despite the large number of incidences recorded. As we cannot test the normality assumption, we chose to take caution and work under the assumption that our data distributions might be skewed. \

__SMART Questions__ \

For our SMART (Specific – Measurable – Attainable – Relevant – Time-oriented) questions we explored these ideas: \

•	Is there a relationship between the time of day and the types of crimes committed? \

•	Is there a relationship between the geographical area and the types of crimes committed? \

•	Does a correlation exist between the two broad categories of crimes – property crimes and violent crimes? \

•	Is there a relationship between the weapon used and the type of crime committed? \


__Data Description - Jiaxi__ \
	__Types of variables and summary__ \
	
	
__Summary Statistics  - Rebecca__ \




__Data Description - Jiaxi__
	__Types of variables and summary__
	
	
__Summary Statistics  - Rebecca__


It is clear that total crime incidences were lower in 2020 than in 2021 for every type of crime besides arson and burglary. The beginning of the COVID-19 pandemic restrictions took place during this time, and we suspect that this is a factor as to why crime rates are lower in 2020. Activities and events were cancelled and people were isolating at home which may have caused less crime activity. Overall, total crime increased by 2,840 incidents. \

__By ward and year- Rebecca__ \

Then, in order to examine the data in each of the different locations and compare them to one another, we found the summary statistics. 
```{r, results='markup'}
#import dataset to environment 
crimedata <- read.csv("data/dc-crimes-search-results.csv")

#removing redundant variables
crimedata1 <- crimedata[, c(2,6,9,10,11,12,23,26,27)]

#changing the blank spaces to NAs
crimedata1[crimedata1 == "" | crimedata1 == " "] <- NA

#converting REPORT_DAT from character to date format 
crimedata1$REPORT_DAT <- as.Date(crimedata1$REPORT_DAT, "%m/%d/%Y")

xkablesummary(crimedata1)
```

We then aggregated crime statistics by type of offense and year. 
```{r,results='markup'}
crimedata20 <- subset(crimedata1[crimedata1$YEAR == 2020, ])
crimedata21 <- subset(crimedata1[crimedata1$YEAR == 2021, ])
crimedata22 <- subset(crimedata1[crimedata1$YEAR == 2022, ])

table(crimedata20$offensekey)
table(crimedata21$offensekey)
table(crimedata22$offensekey)
```
Total incidences of crime in DC increased by 5,913 incidences from 2020 to 2021 but are projected to decrease by 3,073 incidences by the end of 2022.

After that, we broke the data down further by each ward.
```{r,results='markup'}
ward1 <- subset(crimedata1[crimedata1$WARD == 1,])
ward2 <- subset(crimedata1[crimedata1$WARD == 2,])
ward3 <- subset(crimedata1[crimedata1$WARD == 3,])
ward4 <- subset(crimedata1[crimedata1$WARD == 4,])
ward5 <- subset(crimedata1[crimedata1$WARD == 5,])
ward6 <- subset(crimedata1[crimedata1$WARD == 6,])
ward7 <- subset(crimedata1[crimedata1$WARD == 7,])
ward8 <- subset(crimedata1[crimedata1$WARD == 8,])

summary1 <- summary(crimedata1)
summary1
table1 <- sumtable(crimedata1)
```
Ward 3 has the lowest totals for all types of offenses, while the second and third highest totals are more spread out by location. Ward 2 has the highest total incidences in DC.  

```{r,results='markup'}
#creating summary tables of the offense data by ward

table(ward1$offensekey)
table(ward2$offensekey)
table(ward3$offensekey)
table(ward4$offensekey)
table(ward5$offensekey)
table(ward6$offensekey)
table(ward7$offensekey)
table(ward8$offensekey)
```

Across each type of crime, Ward 3 had the lowest amount of incidences. Ward 2 has the highest total amount of crime with Ward 2 following. \

Ward 1 experienced a total increase of 1,337 crimes from 2020 to 2021. Property theft is the most frequent crime while property theft from an automobile saw the largest increase. \

Ward 2 experienced a total increase of 1,875 crimes from 2020 to 2021. Similar to Ward 1, property theft is the most frequent crime while property theft from an automobile saw the largest increase. Both burglary and arson decreased from 2020 to 2021. \

Ward 3 experienced a total increase of only 201 crimes from 2020 to 2021. Burglary and motor vehicle theft decreased by 41 and 44, respectively, but property theft increased by 229. Violent homicide also decreased by 1. \

Ward 4 experienced a total increase of 448 crimes from 2020 to 2021. Violent robbery decreased by 2 while property theft from an automobile was the most frequent crime and had the largest increase. \

Ward 5 experienced a total increase of 829 crimes from 2020 to 2021. Property theft was the most frequent crime in 2020 and property theft from an automobile increased the most and was the most frequent crime in 2021. Arson decreased by 3. \

Ward 6 experienced a total increase of 611 crimes from 2020 to 2021. Property theft was the most frequent crime in both years and experienced the largest increase as well. Arson and violent homicide decreased. \

Ward 7 experienced an increase of only 129 crimes from 2020 to 2021. Property theft from an automobile was the most frequent crime in 2020 but saw the largest decrease, while property theft saw the largest increase and was the most frequent crime in 2021. \

Ward 8 experienced an increase of 456 crimes from 2020 to 2021. Property theft was the most frequent crime and also saw the largest increase in incidences in this area. 

__By time of day - Jiaxi__ \
	
__Correlation coefficient - Jiaxi__ \
	
	
__What does it all mean/Discussion - everybody writes small summaries of their parts and we compile it together__ \

All wards saw an increase in crime from 2020 to 2021. Property theft and theft from a motor vehicle are the types of offenses that are most frequent in all locations in DC. Arson is the least committed type of crime. Overall, crime has increased significantly from 2020 to 2021. This is most likely due to the fact that people were home more frequently because of COVID-19, and then crime activity increased when restrictions were lifted in 2021. 

__Disclosure on Bias- Rebecca__ \

Because crime is self-reported, this database will not include incidences of crime that are not reported or were not humored by law enforcement. Some crimes may be more or less likely to be reported to law enforcement due to societal pressures, stigma, etc. Because of this potential for underrepresentation, this data may inaccurately reflect the true parameters of crime in DC. \

__Future Research - Rebecca__ \

Further research on this topic should address other potentially relevant variables that may help understand incidences of crime in DC. Variables that may be of interest and should be obtained for each ward include population, demographics, income distribution, age distribution, and unemployment levels. \

	__By time of day - Jiaxi__
	
	__Correlation coefficient - Jiaxi__
	
	__Hypothesis testing - Sophia__
	
__What does it all mean/Discussion - everybody writes small summaries of their parts and we compile it together__



__Disclosure on Bias- Rebecca__ \

Because crime is self-reported, this database will not include incidences of crime that are not reported or were not humored by law enforcement. Some crimes may be more or less likely to be reported to law enforcement due to societal pressures, stigma, etc. Because of this potential for under representation, this data may inaccurately reflect the true parameters of crime in DC. \

__Future Research - Rebecca__ \

Further research on this topic should address other potentially relevant variables that may help understand incidences of crime in DC. Variables that may be of interest and should be obtained for each ward include population, demographics, income distribution, age distribution, and unemployment levels. \

__Bibliography__ \

Burdett, Lagos, R., & Wright, R. (2003). Crime, Inequality, and Unemployment. The American Economic Review, 93(5), 1764–1777. https://doi.org/10.1257/000282803322655536

Edmark. (2005). Unemployment and Crime: Is There a Connection? The Scandinavian Journal of Economics, 107(2), 353–373. https://doi.org/10.1111/j.1467-9442.2005.00412.x

Frederick, & Jozefowicz, J. J. (2018). Rural-Urban Differences in the Unemployment-Crime Relationship: The Case of Pennsylvania. Atlantic Economic Journal, 46(2), 189–201. https://doi.org/10.1007/s11293-018-9581-y

  Janko, & Popli, G. (2015). Examining the link between crime and unemployment: A time-series analysis for Canada. Applied Economics, 47(37), 4007–4019. https://doi.org/10.1080/00036846.2015.1023942

Lee. (2016). Unemployment and crime: The role of apprehension. European Journal of Law and Economics, 45(1), 59–80. https://doi.org/10.1007/s10657-016-9526-3



