---
title: 'Summary Paper: Crime in DC'
author: "Group 4: Rebecca Felix, Sophia Mir, Jiaxi Liu"
subtitle: 'DATS 6101 - Spring 2022'
# date: "today"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 

#all libraries go here
library(ezids)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(data.table)
library(vtable)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

__Introduction - Sophia__ \


__Theoretical Framework and Data Source- Sophia__ \


__Data Description - Jiaxi__ \
	__Types of variables and summary__ \

Our dataset covers the crime data from January 2020 to February 2022, with 50,765 observations in total. There are 29 variables in the dataset, 9 of which are used in our project. We removed unuseful data to make sure that our dataset is clean enough. Most of the variables we used are categorical, including offense group, offense type, method, shift, report date, year, and ward. The offense group is used for categorizing offenses into two groups - property or violent crimes. The offense type is subcategories of crimes, including theft/other, theft f/auto, sex abuse, robbery, motor vehicle theft, homicide, burglary, assault w/dangerous weapon, and arson. The method includes three values, which are gun, knife, and other. The shift is a time identifier, which divides the whole day into three parts - day (8 AM to 4 PM), evening (4 PM to 12 AM), and midnight (12 AM to 8 AM). This variable can be different because of seasons. The report date is the time and day the crime was reported to the police department. Similarly, the year is the year the crime was reported. The ward is a location identifier to indicate the specific area the crime happened. There are 2 more variables in our dataset, which are offense text and offense key. These 2 variables are the same with offense type. 
In 2020, there were 22426 crimes happening in the DC area. In 2021, the total number became 28339, there was an increase of 2840 in total. A possible reason can be lockdown because of Covid-19 in 2020. From the total number, we can see that arson was the lowest frequent crime, while theft was the highest frequent crime in both years. 70.5% increase will be caused by the increase in theft crime.\

	
__Summary Statistics  - Rebecca__ \

__Introduction - Sophia__


__Theoretical Framework and Data Source- Sophia__


__Data Description - Jiaxi__
	__Types of variables and summary__
	
	
__Summary Statistics  - Rebecca__


It is clear that total crime incidences were lower in 2020 than in 2021 for every type of crime besides arson and burglary. The beginning of the COVID-19 pandemic restrictions took place during this time, and we suspect that this is a factor as to why crime rates are lower in 2020. Activities and events were cancelled and people were isolating at home which may have caused less crime activity. Overall, total crime increased by 2,840 incidents. \

__By ward and year- Rebecca__ \

Then, in order to examine the data in each of the different locations and compare them to one another, we found the summary statistics. 
```{r, results='markup'}
#import dataset to environment 
crimedata <- read.csv("data/dc-crimes-search-results.csv")

#removing redundant variables
crimedata1 <- crimedata[, c(2,6,9,10,11,12,23,26,27)]

#changing the blank spaces to NAs
crimedata1[crimedata1 == "" | crimedata1 == " "] <- NA

#converting REPORT_DAT from character to date format 
crimedata1$REPORT_DAT <- as.Date(crimedata1$REPORT_DAT, "%m/%d/%Y")

xkablesummary(crimedata1)
```

We then aggregated crime statistics by type of offense and year. 
```{r,results='markup'}
crimedata20 <- subset(crimedata1[crimedata1$YEAR == 2020, ])
crimedata21 <- subset(crimedata1[crimedata1$YEAR == 2021, ])
crimedata22 <- subset(crimedata1[crimedata1$YEAR == 2022, ])

table(crimedata20$offensekey)
table(crimedata21$offensekey)
table(crimedata22$offensekey)
```
Total incidences of crime in DC increased by 5,913 incidences from 2020 to 2021 but are projected to decrease by 3,073 incidences by the end of 2022.

After that, we broke the data down further by each ward.
```{r,results='markup'}
ward1 <- subset(crimedata1[crimedata1$WARD == 1,])
ward2 <- subset(crimedata1[crimedata1$WARD == 2,])
ward3 <- subset(crimedata1[crimedata1$WARD == 3,])
ward4 <- subset(crimedata1[crimedata1$WARD == 4,])
ward5 <- subset(crimedata1[crimedata1$WARD == 5,])
ward6 <- subset(crimedata1[crimedata1$WARD == 6,])
ward7 <- subset(crimedata1[crimedata1$WARD == 7,])
ward8 <- subset(crimedata1[crimedata1$WARD == 8,])

summary1 <- summary(crimedata1)
summary1
table1 <- sumtable(crimedata1)
```
Ward 3 has the lowest totals for all types of offenses, while the second and third highest totals are more spread out by location. Ward 2 has the highest total incidences in DC.  

```{r,results='markup'}
#creating summary tables of the offense data by ward

table(ward1$offensekey)
table(ward2$offensekey)
table(ward3$offensekey)
table(ward4$offensekey)
table(ward5$offensekey)
table(ward6$offensekey)
table(ward7$offensekey)
table(ward8$offensekey)
```

Across each type of crime, Ward 3 had the lowest amount of incidences. Ward 2 has the highest total amount of crime with Ward 2 following. \

Ward 1 experienced a total increase of 1,337 crimes from 2020 to 2021. Property theft is the most frequent crime while property theft from an automobile saw the largest increase. \

Ward 2 experienced a total increase of 1,875 crimes from 2020 to 2021. Similar to Ward 1, property theft is the most frequent crime while property theft from an automobile saw the largest increase. Both burglary and arson decreased from 2020 to 2021. \

Ward 3 experienced a total increase of only 201 crimes from 2020 to 2021. Burglary and motor vehicle theft decreased by 41 and 44, respectively, but property theft increased by 229. Violent homicide also decreased by 1. \

Ward 4 experienced a total increase of 448 crimes from 2020 to 2021. Violent robbery decreased by 2 while property theft from an automobile was the most frequent crime and had the largest increase. \

Ward 5 experienced a total increase of 829 crimes from 2020 to 2021. Property theft was the most frequent crime in 2020 and property theft from an automobile increased the most and was the most frequent crime in 2021. Arson decreased by 3. \

Ward 6 experienced a total increase of 611 crimes from 2020 to 2021. Property theft was the most frequent crime in both years and experienced the largest increase as well. Arson and violent homicide decreased. \

Ward 7 experienced an increase of only 129 crimes from 2020 to 2021. Property theft from an automobile was the most frequent crime in 2020 but saw the largest decrease, while property theft saw the largest increase and was the most frequent crime in 2021. \

Ward 8 experienced an increase of 456 crimes from 2020 to 2021. Property theft was the most frequent crime and also saw the largest increase in incidences in this area. 

__By time of day - Jiaxi__ \
	
__Correlation coefficient - Jiaxi__ \
	
__Hypothesis testing - Sophia__ \
	
__What does it all mean/Discussion - everybody writes small summaries of their parts and we compile it together__ \

All wards saw an increase in crime from 2020 to 2021. Property theft and theft from a motor vehicle are the types of offenses that are most frequent in all locations in DC. Arson is the least committed type of crime. Overall, crime has increased significantly from 2020 to 2021. This is most likely due to the fact that people were home more frequently because of COVID-19, and then crime activity increased when restrictions were lifted in 2021. 

__Disclosure on Bias- Rebecca__ \

Because crime is self-reported, this database will not include incidences of crime that are not reported or were not humored by law enforcement. Some crimes may be more or less likely to be reported to law enforcement due to societal pressures, stigma, etc. Because of this potential for underrepresentation, this data may inaccurately reflect the true parameters of crime in DC. \

__Future Research - Rebecca__ \

Further research on this topic should address other potentially relevant variables that may help understand incidences of crime in DC. Variables that may be of interest and should be obtained for each ward include population, demographics, income distribution, age distribution, and unemployment levels. \

	__By time of day - Jiaxi__
	
	__Correlation coefficient - Jiaxi__
	
	__Hypothesis testing - Sophia__
	
__What does it all mean/Discussion - everybody writes small summaries of their parts and we compile it together__



__Disclosure on Bias- Rebecca__ \

Because crime is self-reported, this database will not include incidences of crime that are not reported or were not humored by law enforcement. Some crimes may be more or less likely to be reported to law enforcement due to societal pressures, stigma, etc. Because of this potential for under representation, this data may inaccurately reflect the true parameters of crime in DC. \

__Future Research - Rebecca__ \

Further research on this topic should address other potentially relevant variables that may help understand incidences of crime in DC. Variables that may be of interest and should be obtained for each ward include population, demographics, income distribution, age distribution, and unemployment levels. \



